General purpose gene annotation using deep learning

### 2020-04-20 paper/main.tex
minor
rewrite approach

\section{Approach}
Helixer is a prototype software for training and utilizing a general purpose model for the functional annotation of large eukaryotic genomes. We demonstrate the effectiveness of this approach by training a single model for animals (phylogenetic names?) and plants respectively, which was then evaluated on 192 genomes in the animal case and on 60 plant genomes. We use modelling techniques from the field of deep learning, which has been shown to have the potential to be superior in flexibility and accuracy to traditional machine learning methods in a variety of fields. Our model expects the raw genomic sequence as main input and produces functional predictions on a base pair wise level. The source code and all data is publically available.

@@

\begin{figure}[!tpb]
\label{fig:length_wise_bias}
\centerline{\includegraphics[width=0.45\textwidth]{images/length_wise_bias}}
\caption{Average Genic F1 and basepair wise accuracy with respect to the position in the 20000 basepair long input sequences of all animals on top and all plant genomes below shown with and without overlapping. Each value is the average performance of a 200 basepair long chunk. The regular metrics show a significant negative bias towards the borders of each sequence. Overlapping the predictions helps to mitigate a lot of that while showing no negative side effects elsewhere on average. See Section \ref{sec:overlapping} for a breakout of the effect of overlapping by each individual species.}
\end{figure}

>>>>

\section{Approach}
Helixer is a prototype software for training and utilizing a general purpose model for the functional cross-species annotation of large eukaryotic genomes using only DNA data as input. We demonstrate the effectiveness of this approach by training a single model for the annotation of a large set of genomes from the kingdoms metazoa and viridiplantae, respectively, which we will call {\it animals} and {\it plants} from now on. In total we used 192 animal genomes and 60 plant genomes. In contrast to traditional approaches, we employ modelling techniques from the field of deep learning to improve both the annotation accuracy and the ability to generalize across genomes. The source code and all data is publically available.

@@

\begin{figure}[!tpb]
\label{fig:length_wise_bias}
\centerline{\includegraphics[width=0.45\textwidth]{images/length_wise_bias}}
\caption{Average Genic F1 and basepair wise accuracy with respect to the position in the 20000 basepair long input sequences of all animals on top and all plant genomes below shown with and without overlapping. Each value is the average performance of a 200 basepair long chunk by one of the final ensemble models. The regular metrics show a significant negative bias towards the borders of each sequence. Overlapping the predictions helps to mitigate a lot of that while showing no negative side effects elsewhere on average. See Section \ref{sec:overlapping} for a breakout of the effect of overlapping by each individual species.}
\end{figure}
